\documentclass[twoside]{article}

\include{packages}
\include{macros}


%\usepackage{aistats2019}
% If your paper is accepted, change the options for the package
% aistats2019 as follows:
%
\usepackage[accepted]{aistats2019}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
\runningauthor{J. Pacheco and J.W. Fisher, III}

\twocolumn[

  \aistatstitle{Variational Information Planning for Sequential Decision Making} 
  \aistatsauthor{ Jason Pacheco \And John W. Fisher, III}
  \aistatsaddress{ MIT CSAIL \And MIT CSAIL} ]

\begin{abstract}  
  We consider the setting of sequential decision making where, at each
  stage, potential actions are evaluated based on expected reduction
  in posterior uncertainty, given by mutual information (MI).  As MI
  typically lacks a closed form, we propose an approach which
  maintains variational approximations of, both, the posterior and MI
  utility.  Our planning objective extends an established variational
  bound on MI to the setting of sequential planning.  The result,
  variational information planning (VIP), is an efficient method for
  sequential decision making.  We further establish convexity of the
  variational planning objective and, under conditional exponential
  family approximations, we show that the optimal MI bound arises from
  a relaxation of the well-known exponential family moment matching
  property.  We demonstrate VIP for sensor selection, experiment
  design, and active learning, where it meets or exceeds methods
  requiring more computation, or those specialized to the 
  task.
\end{abstract}

\section{Introduction}
\input{introduction}

\section{Sequential Information Planning}
\input{infoplan}

\section{Variational Information Planning}\label{sec:varplan}
\input{varplan_content}

\section{Optimization for Conditional Exponential Families}\label{sec:optim}
\input{optim}

\section{Evaluating the MI Bound}\label{sec:eval}
\input{expfamapprox}

\section{Experimental Results}\label{sec:experiments}
\input{experiments}

\section{Conclusion}
\input{conclusion}

\textbf{Acknowledgments} The first author would like to thank Sue
Zheng for useful technical discussions.  This work was partially
supported by the ONR (N00014-17-1-2072), the Department of Energy
(CVT Consortium), and DNDO under the ARI program.

\bibliography{refs}
\bibliographystyle{abbrvnat}

\end{document}
